---
title: "Lab 12 Homework"
author: "Luc-Tanton Tran"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above. For any included plots, make sure they are clearly labeled. You are free to use any plot type that you feel best communicates the results of your analysis.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
``` 

```{r}
options(scipen = 999)
```

## Resources
The idea for this assignment came from [Rebecca Barter's](http://www.rebeccabarter.com/blog/2017-11-17-ggplot2_tutorial/) ggplot tutorial so if you get stuck this is a good place to have a look.  

## Gapminder
For this assignment, we are going to use the dataset [gapminder](https://cran.r-project.org/web/packages/gapminder/index.html). Gapminder includes information about economics, population, and life expectancy from countries all over the world. You will need to install it before use.  
```{r}
#install.packages("gapminder")
library("gapminder")
```

1. Use the function(s) of your choice to get an idea of the overall structure of the data frame, including its dimensions, column names, variable classes, etc. As part of this, determine how NA's are treated in the data.   

*There are 6 columns and 1,704 rows. There are continuous variables, factors, and integers (year). Year may have to be reclassified as a factor so that the bar plots look good. Off cursory glance, there doesn't seem to be any obvious means of categorizing NAs.*
```{r}
gapminder <- gapminder
glimpse(gapminder)
summary(gapminder)
```

2. How many countries are represented in this dataset? Make a table and a plot that shows the number of countries by continent.  

*There are 142 total countries.There are 52 countries in Africa, 25 in the Americas, 33 in Asia, 30 in Europe, and 2 in Oceania.*
```{r}
gapminder %>% 
  summarize(n_countries = n_distinct(country))

gapminder %>% 
  group_by(continent) %>% 
  summarize(n_countries = n_distinct(country))

gapminder %>% 
  ggplot(aes(x=continent))+
  geom_bar()+
  labs(title = "Number of Countries by Continent",
       x = "Continent",
       y = "Number of Countries")+
  theme(plot.title = element_text(size = rel(2), hjust = 0.5))
  
```

3. Which country has the largest population growth since 1952? Show this as a table.  
```{r}
gapminder %>% 
  group_by(country) %>% 
  summarize(
    pop_growth = max(pop)-min(pop),
    log10_pop_growth = log(pop_growth), #for analysis
    .groups='keep'
  ) %>% 
  arrange(-pop_growth)
```


4. Make a plot that shows population growth for the country you found in question #3. This plot should show the change over time.  
```{r}
gapminder %>% 
  filter(country == "China") %>% 
  ggplot(aes(x=year, y = pop))+
  geom_point()+
  geom_line()+
  labs(
    title = "Population of China Since 1952 on a log scale",
    x = "Year",
    y = "Population"
  )+
  scale_y_log10()+ #scaling population to log10
  theme(plot.title = element_text(hjust = 0.5))
```

5. How has global life expectancy changed between 1952 and 2007? Show the min, mean, and max for all countries in the dataset. Show this as a table.  
```{r}
# Min, Mean, Max for each country
gapminder %>% 
  group_by(country) %>% 
  summarize(
    min_expectancy = min(lifeExp),
    mean_expectancy = mean(lifeExp),
    max_expectancy = max(lifeExp)
  )
```

6. Make a plot that shows how mean life expectancy has changed over time for each continent. What is your interpretation of what happened in Africa between 1987 and 2002?  

*It looks like something happened during that time. From quick googling, it seems that it might be from floods at the time and relatively inaccessible medical services. There's also a chance it's because of AIDS.*
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  mutate(mean_life_exp = mean(lifeExp)) %>% 
  ggplot(aes(x = year, y = mean_life_exp, color = continent))+
  geom_point()+
  geom_line()+
  labs(title = "Mean Life Expectancy Over Time by Continent",
       x = "Year",
       y = "Average Life Expectancy")+
  theme(plot.title = element_text(hjust=0.5))
```

7. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer? Show this as a plot.  

*It looks like it helps to a degree, because there's a general upward trend. It's very noisy, however.*
```{r}
gapminder %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp))+
  geom_point()+
  geom_line()+
  scale_x_log10()+
  labs(title = "Life expectancy vs GDP Per Capita with GDP on a log scale",
       x = "GDP Per Capita",
       y = "Life Expectancy")+
  theme(plot.title = element_text(hjust=0.5))
```

8. Which five countries have had the highest GDP per capita growth over the years represented in this dataset? Show this as a table.  

*Because it's a bit vague on what "highest GDP per capita growth over the years represented", I'm interpreting this question as "the biggest change if you take the first and last values and take the difference."*

*Using this definition, the countries are Singapore, Norway, Hong Kong (China), Ireland, and Austria.*
```{r}
gapminder %>% 
  group_by(country) %>% 
  summarize(
    first_gdp = first(gdpPercap, order_by = year),
    last_gdp = last(gdpPercap, order_by = year),
    gdp_diff = last_gdp - first_gdp
  ) %>% 
  top_n(5) %>% 
  arrange(-gdp_diff)
```

9. How does per capita GDP growth compare between these same five countries? Show this as a plot.  
```{r}
gapminder %>% 
  filter(country %in% c("Singapore", "Norway", "Hong Kong, China", "Ireland", "Austria")) %>% 
  ggplot(aes(x = year, y = gdpPercap, color = country))+
  geom_point()+
  geom_line()+
  labs(
    title = "GDP Per Capita vs Year by Country",
    x = "Year",
    y = "GDP Per Capita"
  )+
  theme(plot.title = element_text(hjust= 0.5))
```

10. Do one analysis of your choice that includes a table and plot as outputs.  

**Convert** `gdpPercap` **into categorical variables, and make a table of each year and how many countries fall into each category. Then, graph number of countries (y-axis) vs year (x-axis) grouped by each gdp category. What trends do you notice?**

*I notice that there seems to be a general uptick in countries with higher GDPs. Obviously, there will be some that aren't in the highest quartiles, but in the overall data, it seems like it's approaching roughly 42% of countries doing in the top 25% of all data (~60/142 countries, "high" gdp being the third quartile or higher).*

```{r}
library(gtools)
quartiles <- quantcut(gapminder$gdpPercap)
table(quartiles)

gapminder <- gapminder %>% 
  mutate(gdp_category = case_when(gdpPercap < 1.2e+03 ~ "low",
                                  gdpPercap > 1.2e+03 & gdpPercap <= 3.53e+03 ~ "medium-low",
                                  gdpPercap > 3.53e+03 & gdpPercap <= 9.33e+03 ~ "medium-high",
                                  gdpPercap > 9.33e+03 ~ "high"))

gapminder %>% 
  group_by(gdp_category, year) %>% 
  summarize(
    n_countries = n_distinct(country),
    .groups='keep'
  ) %>% 
  pivot_wider(
    names_from = gdp_category,
    values_from = n_countries
  )

gapminder %>% 
  group_by(year, gdp_category) %>% 
  summarize(n=n(), .groups='keep') %>% 
  ggplot(aes(x=year, y = n, group = gdp_category, color = gdp_category))+
  geom_point()+
  geom_line()+
  labs(title = "How many countries are in each GDP category by year",
       x = "Year",
       y = "Number of Countries")
```

## Knit and Upload
Please knit your work as a .pdf or .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! 