---
title: "Homework 3"
author: "Luc-Tanton Tran"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions

Answer the following questions and/or complete the exercises in RMarkdown. Please embed all of your code and push the final work to your repository. Your report should be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run.

## Load the tidyverse

```{r}
library("tidyverse")
options(scipen=999) # turn off scientific notation
```

## Data

For the homework, we will use data about vertebrate home range sizes. The data are in the class folder and the reference is below.

**Database of vertebrate home range sizes.**. Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. <http://dx.doi.org/10.1086/682070>.\
Data: <http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1>

**1. Load the data into a new object called `homerange`.**

```{r}
homerange <- read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```

**2. What are the dimensions of the dataframe?**

```{r}
dim(homerange)
```

**3. Are there any NA's in the dataframe? Try using summary to determine which variables have more or less NA's.**

```{r}
anyNA(homerange)
summary(homerange)
```

**4. What are the names of the columns in the dataframe?**

```{r}
colnames(homerange)
```

**5. Based on the summary output, do you see anything in the data that looks strange? Think like a biologist and consider the variables.**\
There's so many things. The Genus should *technically* be capitalized in order for it to be capitalized, as per convention of scientific names.

While I was able to figure out from the research paper that the variable `dimension` referred to foraging dimensionality, it is unclear from the name.

There is also a lot of `NA` for the `preymass` variable, which prompted me to check if it was because some are carnivores, but many carnivores also had `NA` counts.

There's also some entries without an `N` value, but are somehow able to calculate `mean.mass.g`, which makes me wonder how they calculated these values.

There's also inconsistency in the variable naming: some use periods to separate words, while some are just shoved together without any such separation.

**6. The `min` and `max` functions can be used to find the minimum and maximum values in a vector. Use these functions to find the minimum and maximum values for the variable `mean.mass.g`.**

```{r}
min(homerange$mean.mass.g)
max(homerange$mean.mass.g)
```

**7. Change the class of the variables `taxon` and `order` to factors and display their levels.**

```{r}
homerange$taxon <- as.factor(homerange$taxon)
levels(homerange$taxon)
homerange$order <- as.factor(homerange$order)
levels(homerange$order)
```

**8. Use `select` to pull out the variables taxon and common.name.**

```{r}
select(homerange, taxon, common.name)
```

**9. Use `filter` to pull out all mammals from the data.**

```{r}
filter(homerange, taxon == "mammals")
```

**10. What is the largest `mean.hra.m2` for mammals? This is a very large number! Which animal has this homerange? Look them up online and see if you can figure out why this number is so large.**

```{r}
mammals <- filter(homerange, taxon == "mammals") 
max(mammals$mean.hra.m2)
filter(mammals, mean.hra.m2 == max(mammals$mean.hra.m2))
```

From what I can find online, they need to travel a large distance (and are found in many places), and this large number reflects that.

## Knit and Upload

Please knit your work as a .pdf or .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!
